---
title: "Metabolome paper - Calculate microbiome beta NTI"
Author: "Hannah Holland Moritz"
output:
  html_document: default
  pdf_document: default
date: "2023-03-26"
editor_options:
  chunk_output_type: console
---

#' ## Calculate betaNTI
#' This step calcluates the betaNTI or phylogenetic turnover among samples
#' It requires a phylogenetic tree and a species-by-sample table
#' This step takes a long time and would be much better run on a super computer
#+ include=FALSE

# Loading libraries
```{r}
# Load DADA2 and required packages
library(tidyverse); packageVersion("tidyverse") # for dataframe processing
library(picante); packageVersion("picante") # handles tree manipulations
library(vegan); packageVersion("vegan") # for ecological applications
library(here)
library(doMPI)

```

# Load required data
```{r}
source(here("setup.R")) 
```

## Set up input and output directories
#+ directory creation, eval = FALSE
# setting up names of file paths to stay organized

```{r}
outputs.fp <- here("Assembly-analysis", "outputs")
figures.fp <- here("Assembly-analysis", "figures")

if (!dir.exists(outputs.fp)) {dir.create(outputs.fp)}
if (!dir.exists(figures.fp)) {dir.create(figures.fp)}

```

#' First we'll set up user settings and command line arguments: 
#' Users should modify these settings for their own environment.  
#'  - Set the number of replicate randomizations for the BetaMNTD calculations `beta.reps`
#'  - Set the relative abundance flag if your OTU table is not integers but real numbers instead
#'  - If you're working in an environment with parallel computing set `paral <- TRUE`
# User settings and command line arguments

```{r}
beta.reps <- 1000; # number of randomizations for beta MNTD

rel_abund <- TRUE # is the otu table in relative abundances such as read mappings?

paral <- TRUE

# Tell the user their settings:
writeLines(paste("User settings are as follows: \n",
                 "Parallel:", paral, "\n",
                 "Replicates:", beta.reps, "\n",
                 "Relative Abundances:", rel_abund))
# If parallel is true, detect available cores

if(paral==T) {
  library(Rmpi)  # R implementation of MPI interface
  library(doMPI)
  n.cores <- max(1, mpi.universe.size()-1) # set n.cores to 1 or mpi allocation size, whichever is bigger
  #print(n.cores)
  print(paste(n.cores, "worker nodes are available."))
  
  if (n.cores > 24) {
    # we are using more than 1 node, so really run in parallel mode
    cl <- startMPIcluster(count = n.cores)
    registerDoMPI(cl) # tell foreach about the cluster
    print(paste("Running in parallel mode on",n.cores,"worker nodes."))
    
  } else if (n.cores > 1) {
    library(doParallel)
    cl <- parallel::makeForkCluster(n.cores)
    registerDoParallel(cl)
    print(paste("Running in parallel mode without mpi on", n.cores, "worker nodes."))
  } else {
    registerDoSEQ() # tells foreach to use sequential mode
    paral <- FALSE
    print(paste("Only", n.cores,
                "available. Not enough to run in parallel, running in sequential mode."))
  }
} else {
  registerDoSEQ()
  print("Running in sequential mode.")
} 
```


## Read in data 

```{r}

writeLines("reading in input data")
input <- input_ra

# Read in tree
magtree <- input_ra$tree

# Verify the everything is matching
match.phylo.otu <- match.phylo.data(magtree, input_ra$otu_table[-1])

```

#' ## Step 1: First we calculate the empirical betaMNTD
#' This is the observed betaMNTD which we will compare to our null distribution that we will generate below

```{r}

beta.mntd.weighted <- as.matrix(comdistnt(t(match.phylo.otu$data),
                                          cophenetic(match.phylo.otu$phy),
                                          abundance.weighted=T))
dim(beta.mntd.weighted)

```

#' ## Step 2: Calculate null expectation (randomized) betaMNTD
#' This is our null distribution. Calculating this will take a while - best to run on server
```{r}
null_bMNTD <- NULL # make a null object to hold list of results
null_bMNTD <- foreach (i=1:beta.reps, .packages = c("picante")) %dopar% {
  as.matrix(comdistnt(t(match.phylo.otu$data),
                      taxaShuffle(cophenetic(match.phylo.otu$phy)),
                      abundance.weighted=T,
                      exclude.conspecifics = F))
  } # end of parallel computation
```

# reformat data output from foreach
# Unlisting the list of null beta.reps; This creates an 3d array
# of size: # samples x # samples x # beta.reps

```{r}
rand.weighted.bMNTD.comp <- array(as.numeric(unlist(null_bMNTD)), 
                                  dim=c(ncol(match.phylo.otu$data),
                                        ncol(match.phylo.otu$data), 
                                        beta.reps))

dim(rand.weighted.bMNTD.comp) # should be # samples x # samples x # beta.reps
```

#' ## Step 3: Now calculate betaNTI by normalizing observed against the null expectation
#' Reminder: 
#' -2 < bNTI < 2 indicates stochastic process which can be further refined using values from RCbray in next step.
#' bNTI < -2 indicates homogeneous selection
#' bNTI > 2 indicates heterogeneous selection
```{r}
# Save time by setting up matrix first
weighted.bNTI <- matrix(c(NA),nrow=ncol(match.phylo.otu$data),ncol=ncol(match.phylo.otu$data));

dim(weighted.bNTI); # should be # samples x # samples

for (columns in 1:(ncol(match.phylo.otu$data)-1)) {
  for (rows in (columns+1):ncol(match.phylo.otu$data)) { # columns + 1 so samples are not compared to themselves
    
    # Pull out the randomized betaMNTD values
    rand.vals <- rand.weighted.bMNTD.comp[rows,columns,]; # length of this object = beta.reps
    # Normalize the observed against the randomized values
    weighted.bNTI[rows,columns] <- (beta.mntd.weighted[rows,columns] - mean(rand.vals)) / sd(rand.vals); # normalize
    rm("rand.vals");
    
  };
};

# Rename columns and rows to get a samplexsample table
rownames(weighted.bNTI) = colnames(match.phylo.otu$data);
colnames(weighted.bNTI) = colnames(match.phylo.otu$data)
# weighted.bNTI;
```

## Save Data

```{r}
# BetaNTI Matrix
write.csv(weighted.bNTI,
          paste0(outputs.fp, "/weighted_BNTI_bacteria.csv"),
          quote=FALSE)
```

