---
title: "Metabolome paper - Figure 1"
output:
  html_document: default
  pdf_document: default
date: "2023-03-26"
editor_options:
  chunk_output_type: console
---
# Loading libraries
```{r}
library(reshape2)
require(vegan)
library(ggsci)
library(rstatix)
library(ggpubr)
library(ggrepel)
library(grid)
library(gridExtra)
library(phyloseq)
library(ggridges)
library(patchwork)
library(corrplot)
library(tidyverse)
```


# Functions

```{r}
distance_func <- function(bnti_matrix, geo_matrix, geo_col){
  
  geo_selected <- geo_matrix %>% 
    select(geo_col, SampleID) %>% 
    rename(Selected = geo_col) %>% 
    filter(!is.na(Selected))
  
  bnti_filt <- bnti_matrix[rownames(bnti_matrix) %in% geo_selected$SampleID, 
                           colnames(bnti_matrix) %in% geo_selected$SampleID]
  
  geo_selected <- geo_selected %>% 
    filter(SampleID %in% rownames(bnti_filt))
  
  bnti_filt <- bnti_filt[geo_selected$SampleID, geo_selected$SampleID]
  
  n <- length(rownames(bnti_filt))
  dist_mat <- as.data.frame(matrix(nrow = n, ncol = n))
  rownames(dist_mat) <- rownames(bnti_filt)
  colnames(dist_mat) <- colnames(bnti_filt)
  
  for(i in 1:(n-1)){
    for(j in (i+1):n){
      dist_mat[j, i] <- abs(geo_selected$Selected[j] - geo_selected$Selected[i])
    }
  }
  dist_mat[upper.tri(dist_mat)] = t(dist_mat)[upper.tri(dist_mat)]
  return(dist_mat)
  
}
```


```{r}
mantel_fun <- function(bnti_matrix, geo_matrix){
  
  bnti_filt <- bnti_matrix[rownames(bnti_matrix) %in% rownames(geo_matrix), 
                           colnames(bnti_matrix) %in% colnames(geo_matrix)]
  
  mantel_test <- mantel(bnti_filt, geo_matrix, 
                        method = "pearson", permutations = 9999)
  
  return(mantel_test)
}

```


# Plots colors
```{r}
list_colors <- c('Palsa' = '#703C1B', 'Bog' = '#058000','Fen' = '#0001FF')
```

# Figure 1

## Violin metabolites

### Loading BNTI matrix and data preprocessing

```{r}
# Load in bNTI
weig <- read_csv("../Assembly_Metabolome/output/Abisko_TWCD_bNTI.csv") %>% 
  column_to_rownames(var = "...1")

# Load in RCBC
weig.rcbc <- read_csv("../Assembly_Metabolome/output/Abisko_TWCD_RCBC_9999.csv")%>% 
  column_to_rownames(var = "...1")

## Data Reprocessing

# Matching the order of RCBC to bNTI results
weig.rcbc = weig.rcbc[row.names(weig), colnames(weig), drop = F]

# Setting the RCBC diagonals to NA

diag(weig.rcbc) = NA

# Reflecting null matrices

weig[upper.tri(weig)] = t(weig)[upper.tri(weig)]

# Removing significant bNTI results from the RCBC results

weig.rcbc[abs(weig) > 2] = NA

```

### Preprocessing Data for plotting

```{r}
# Melting data
weig = melt(as.matrix(weig)); weig$Type = "Bulk"

weig.rcbc = melt(as.matrix(weig.rcbc)); weig.rcbc$Type = "All"

#Combine data and remove null values

bnti_data <- weig %>% 
  filter(!is.na(value))

rcbc_data <- weig.rcbc %>% 
  filter(!is.na(value))

# Adding Habitat Information
bnti_data$Habitat = "Bog"
bnti_data$Habitat[grep("E", bnti_data$Var2)] = "Fen"
bnti_data$Habitat[grep("P", bnti_data$Var2)] = "Palsa"

bnti_data$Habitat <- factor(bnti_data$Habitat, levels = c("Palsa", "Bog", "Fen"))

rcbc_data$Habitat = "Bog"
rcbc_data$Habitat[grep("E", rcbc_data$Var2)] = "Fen"
rcbc_data$Habitat[grep("P", rcbc_data$Var2)] = "Palsa"

rcbc_data$Habitat <- factor(rcbc_data$Habitat, levels = c("Palsa", "Bog", "Fen"))

```

### BNTI plot all metabolites

```{r}

# Boxplots by like samples (example: Palsa vs Palsa)

bnti_data_within <- bnti_data %>% 
  filter(
    (str_detect(Var1, '_P_') & str_detect(Var2, '_P_')) |
      (str_detect(Var1, '_E_') & str_detect(Var2, '_E_')) |
      (str_detect(Var1, '[a-z]_S_') & str_detect(Var2, '[a-z]_S_'))
  ) %>% 
  mutate(Type = str_replace(Type, 'All', 'Bulk'))


stat_df <- bnti_data_within %>% 
  wilcox_test(value~Habitat) %>% 
  adjust_pvalue(method = 'bonferroni') %>% 
  add_significance() %>%
  add_xy_position()

#Violin plot

bp_group_vio_so <- bnti_data_within %>% 
  mutate(Type = "Bulk metabolites") %>% 
  ggplot(aes(x = Habitat, y = value, fill = Habitat))+
  geom_boxplot(width = 0.2, show.legend = FALSE) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_hline(yintercept = c(-2,2), color = "red", lty = 2)+
  labs(#title = expression(bold(paste(beta, "NTI ", metabolites))),
       y = expression(bold(paste(beta, 'NTI'))),
       fill = 'Habitat')+
  scale_fill_manual(values = list_colors)+
  stat_pvalue_manual(stat_df,
                     label = 'p.adj.signif', inherit.aes = FALSE, hide.ns = TRUE)+
  facet_wrap(~Type)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        #axis.title.y = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_rect(fill = 'black'),
        strip.text = element_text(color = 'white', face = 'bold'))


bp_group_vio_so
```

## Violing microbes

### Load data

```{r}
## Bacteria BNTI

samples_names <- read_csv("../Assembly_Microbiome/data/metadata_matched_MAGs_data_11-29-21.csv")

bacteria_bnti <- read_csv("../Assembly_Microbiome/weighted_BNTI_bacteria.csv") %>% 
  column_to_rownames("...1")

bacteria_bnti <- bacteria_bnti[colnames(bacteria_bnti),] %>% 
  rename(!!(set_names(samples_names$names_bnti, samples_names$SampleID))) 

rownames(bacteria_bnti) <- colnames(bacteria_bnti)

# Reflecting  matrices

bacteria_r <- bacteria_bnti
bacteria_r[upper.tri(bacteria_bnti)] <-  t(bacteria_bnti)[upper.tri(bacteria_bnti)]

# Melting data
bnti_data_bacteria = melt(as.matrix(bacteria_r))

#Combine data and remove null values

bnti_data_bacteria <- bnti_data_bacteria %>% 
  filter(!is.na(value))

# Adding Habitat Information
bnti_data_bacteria$Habitat = "Bog"
bnti_data_bacteria$Habitat[grep("E", bnti_data_bacteria$Var2)] = "Fen"
bnti_data_bacteria$Habitat[grep("P", bnti_data_bacteria$Var2)] = "Palsa"

bnti_data_bacteria$Habitat <- factor(bnti_data_bacteria$Habitat, levels = c("Palsa", "Bog", "Fen"))

bnti_data_bacteria <- bnti_data_bacteria %>% 
  mutate(Type = "microbe")
```

### BNTI violin plot

```{r}
# Boxplots by like samples (example: Palsa vs Palsa)

bnti_data_within_bac <- bnti_data_bacteria %>% 
  filter(
    (str_detect(Var1, '_P_') & str_detect(Var2, '_P_')) |
      (str_detect(Var1, '_E_') & str_detect(Var2, '_E_')) |
      (str_detect(Var1, '[a-z]_S_') & str_detect(Var2, '[a-z]_S_'))
  )

stat_df <- bnti_data_within_bac %>% 
  wilcox_test(value~Habitat) %>% 
  adjust_pvalue(method = 'bonferroni') %>% 
  add_significance() %>%
  add_xy_position()

# Boxplots by group
# Violin plot

bp_group_vio_bacteria <- bnti_data_within_bac %>% 
  mutate(type = 'Microbiome') %>% 
  ggplot(aes(x = Habitat, y = value, fill = Habitat))+
  geom_boxplot(width = 0.2, show.legend = FALSE) +
  geom_violin(alpha = 0.5, show.legend = FALSE) +
  geom_hline(yintercept = c(-2,2), color = "red", lty = 2)+
  labs(y = expression(bold(paste(beta, 'NTI'))),
       fill = 'Habitat')+
  scale_fill_manual(values = list_colors)+
  stat_pvalue_manual(stat_df,
                     label = 'p.adj.signif', inherit.aes = FALSE, hide.ns = TRUE)+
  facet_wrap(~type) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(hjust = 0.5),
        # axis.text.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.background = element_rect(fill = 'black'),
        strip.text = element_text(color = 'white', face = 'bold'))
bp_group_vio_bacteria


```


#Combine data and remove null values
```{r}

#Combine data and remove null values

bnti_data_full <- rbind(bnti_data_bacteria, bnti_data) %>% 
  filter(!is.na(value)) %>% 
  mutate(Type = str_replace_all(Type, "All", "Bulk metabolites")) %>% 
  mutate(Type = str_replace_all(Type, "microbe", "Microbiome"))

bnti_data_full$Habitat <- factor(bnti_data_full$Habitat, levels = c("Palsa", "Bog", "Fen"))

```

## Density plots per habitat


```{r}
# Density plots "within" by like samples (example: Palsa vs Palsa)

bnti_data_within_full <- bnti_data_full %>% 
  filter(
    (str_detect(Var1, '_P_') & str_detect(Var2, '_P_')) |
      (str_detect(Var1, '_E_') & str_detect(Var2, '_E_')) |
      (str_detect(Var1, '[a-z]_S_') & str_detect(Var2, '[a-z]_S_')))
```

## Density plots per season

### Density plot by season

### Violin plot with color points colored as season and depth

```{r}
# Filtering month * month 

bnti_data_within_june <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'June') & str_detect(Var2, 'June')) %>% 
  mutate(season = "June")


bnti_data_within_july <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'July') & str_detect(Var2, 'July'))%>% 
  mutate(season = "July")

bnti_data_within_aug <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'Aug') & str_detect(Var2, 'Aug'))%>% 
  mutate(season = "Aug")

bnti_data_season <- rbind(bnti_data_within_june,
                          bnti_data_within_july,
                          bnti_data_within_aug)
bnti_data_season$season  <- factor(bnti_data_season$season, levels = c("June", "July", "Aug"))


```

```{r}
month_den <- bnti_data_season %>% 
  filter(Type == 'Microbiome' | Type == 'Bulk') %>%
  mutate(Type = str_replace(Type, "Bulk", "Bulk metabolites")) %>% 
  mutate(season = factor(season, levels = c('Aug', 'July', 'June'))) %>% 
  mutate(Type = factor(Type, levels = c("Microbiome", "Bulk metabolites"))) %>%
  ggplot(aes(x = value, y = season, fill = Habitat))+
  geom_density_ridges(alpha = 0.5,
                      scale = .8) +
  scale_y_discrete(expand = c(0,0)) +
  geom_vline(xintercept = c(-2,2), color = "red", lty = 2)+
  labs(#title = expression(bold("Density plot")),
       y = expression(bold("Density")),
       x = expression(bold(paste(beta, "NTI"))))+
  scale_fill_manual(values = list_colors)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 15),
        axis.title.y = element_text(hjust = 0.5)) + 
  facet_grid(cols = vars(Type), scales = 'free') +
  theme(strip.background = element_rect(fill = 'black'),
        strip.text = element_text(color = 'white', face = 'bold'))

month_den

```


## Density plots per depth

### Density plot by depth

### Violin plot with color points colored as season and depth

```{r}
# Filtering month * month 

bnti_data_within_surface <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'S_12') & str_detect(Var2, 'S_12')) %>% 
  mutate(depth = "Surface")


bnti_data_within_middle <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'M') & str_detect(Var2, 'M'))%>% 
  mutate(depth = "Middle")

bnti_data_within_deep <- bnti_data_within_full %>%
  filter(str_detect(Var1, 'D') & str_detect(Var2, 'D'))%>% 
  mutate(depth = "Deep")


bnti_data_depth <- rbind(bnti_data_within_surface,
                          bnti_data_within_middle,
                          bnti_data_within_deep)

bnti_data_depth$depth  <- factor(bnti_data_depth$depth, levels = c("Surface", "Middle", "Deep"))


```

```{r}
depth_den <- bnti_data_depth %>% 
  filter(Type == 'Microbiome' | Type == 'Bulk') %>%
  mutate(Type = str_replace(Type, "Bulk", "Bulk metabolites")) %>% 
  mutate(depth = factor(depth, levels = c("Deep", "Middle", "Surface"))) %>% 
  mutate(Type = factor(Type, levels = c("Microbiome", "Bulk metabolites"))) %>% 
  ggplot(aes(x = value, y = depth, fill = Habitat, show_le))+
  geom_density_ridges(alpha = 0.5,
                      scale = .8) +
  scale_y_discrete(expand = c(0,0)) +
  geom_vline(xintercept = c(-2,2), color = "red", lty = 2)+
  labs(#title = expression(bold("Density plot")),
       y = expression(bold("Density")),
       x = expression(bold(paste(beta, "NTI"))))+
  scale_fill_manual(values = list_colors)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = 15),
        axis.title.y = element_text(hjust = 0.5)) + 
  facet_grid(cols = vars(Type), scales = 'free') +
  theme(strip.background = element_rect(fill = 'black'),
        strip.text = element_text(color = 'white', face = 'bold'))

depth_den

```


## Ecological processes

### Colors

```{r}
eco_process_colors <- c('Homogeneous selection' = '#ED553B', 
                        'Variable selection' = '#F9A95A', 
                        'Undominated' = '#3CAEA3', 
                        'Homogenizing dispersal' = '#5894C1' ,
                        'Dispersal limitation' = '#89C6F5')
```

### Loading calculated ecological process and filtering

#### Metabolites

```{r}
data_metabolites <- read.csv("input/Ecological_process_pw_peat_1-25-22.csv") %>% 
  mutate(Compound = str_replace(Compound, "All" , "Bulk metabolites"),
         Habitat = factor(Habitat, levels = c('Palsa', 'Bog', 'Fen'))) %>% 
  select(-Var1, -Type) %>% 
  rename(Assembly_process = Var2) %>% 
  mutate(Assembly_process = str_replace(Assembly_process, 
                                        'Homogenous', 'Homogeneous'))
```

#### Bacteria

```{r}
#Loading RCBC matrix
rcbc_bac <- read_tsv("input/rcbc_matrix_changed_names.tsv") %>% 
  column_to_rownames(var = "...1")

## Data Preprocessing
rcbc_bac <- rcbc_bac[colnames(rcbc_bac),] %>% 
  rename(!!(set_names(samples_names$names_bnti, samples_names$SampleID)))
rownames(rcbc_bac) <- colnames(rcbc_bac)

# Reflecting null matrices
rcbc_r <- rcbc_bac
rcbc_r[upper.tri(rcbc_bac)] <-  t(rcbc_bac)[upper.tri(rcbc_bac)]

# Removing significant bNTI results from the RCBC results
rcbc_r[abs(bacteria_r) > 2] = NA

# Melting data
rcbc_r = melt(as.matrix(rcbc_r))

#Combine data and remove null values
rcbc_data_bac <- rcbc_r %>% 
  filter(!is.na(value)) %>% 
  mutate(Habitat = case_when(str_detect(Var2, 'E') ~ 'Fen',
                             str_detect(Var2, 'P') ~ 'Palsa',
                             TRUE ~ 'Bog'),
         Habitat = factor(Habitat, levels = c('Palsa', 'Bog', 'Fen')))

```

##### Determining ecological processes fractionation

```{r}
#BOG only

bnti_data_bog = bnti_data_bacteria[-c(grep("E", bnti_data_bacteria$Var1), 
                             grep("E", bnti_data_bacteria$Var2),
                             grep("P", bnti_data_bacteria$Var1), 
                             grep("P", bnti_data_bacteria$Var2)),]

rcbc_data_bog = rcbc_data_bac[-c(grep("E", rcbc_data_bac$Var1), 
                             grep("E", rcbc_data_bac$Var2),
                             grep("P", rcbc_data_bac$Var1), 
                             grep("P", rcbc_data_bac$Var2)),]


# Determining ecological processes fractionation
eco.proc_2 = data.frame(hom.sel = NA, 
                        hom.dis = NA, 
                        und = NA, 
                        dis.lim = NA, 
                        var.sel = NA, 
                        row.names = "Bog")

eco.proc_2$hom.sel = length(which(bnti_data_bog$value <= -2)) / 
  length(bnti_data_bog$value)

eco.proc_2$hom.dis = length(which(rcbc_data_bog$value <= -0.95)) / 
  length(bnti_data_bog$value)

eco.proc_2$und = length(which(abs(rcbc_data_bog$value) < 0.95)) / 
  length(bnti_data_bog$value)

eco.proc_2$dis.lim = length(which(rcbc_data_bog$value >= 0.95)) / 
  length(bnti_data_bog$value)

eco.proc_2$var.sel = length(which(bnti_data_bog$value >= 2)) / 
  length(bnti_data_bog$value)

eco.proc_2 = melt(as.matrix(eco.proc_2))

eco.proc_2$value = (eco.proc_2$value*100)

```


```{r}
#FEN only

bnti_data_fen = bnti_data_bacteria[-c(grep("[a-z]_S_", bnti_data_bacteria$Var1), 
                             grep("[a-z]_S_", bnti_data_bacteria$Var2),
                             grep("P", bnti_data_bacteria$Var1), 
                             grep("P", bnti_data_bacteria$Var2)),]

rcbc_data_fen = rcbc_data_bac[-c(grep("[a-z]_S_", rcbc_data_bac$Var1), 
                             grep("[a-z]_S_", rcbc_data_bac$Var2),
                             grep("P", rcbc_data_bac$Var1), 
                             grep("P", rcbc_data_bac$Var2)),]


# Determining ecological processes fractionation
eco.proc_fen = data.frame(hom.sel = NA, 
                          hom.dis = NA, 
                          und = NA, 
                          dis.lim = NA, 
                          var.sel = NA, 
                          row.names = "Fen")

eco.proc_fen$hom.sel = length(which(bnti_data_fen$value <= -2)) / 
  length(bnti_data_fen$value)

eco.proc_fen$hom.dis = length(which(rcbc_data_fen$value <= -0.95)) / 
  length(bnti_data_fen$value)

eco.proc_fen$und = length(which(abs(rcbc_data_fen$value) < 0.95)) / 
  length(bnti_data_fen$value)

eco.proc_fen$dis.lim = length(which(rcbc_data_fen$value >= 0.95)) / 
  length(bnti_data_fen$value)

eco.proc_fen$var.sel = length(which(bnti_data_fen$value >= 2)) / 
  length(bnti_data_fen$value)

eco.proc_fen = melt(as.matrix(eco.proc_fen))

eco.proc_fen$value = (eco.proc_fen$value*100)

```

```{r}
#PALSA only

bnti_data_palsa = bnti_data_bacteria[-c(grep("[a-z]_S_", bnti_data_bacteria$Var1), 
                               grep("[a-z]_S_", bnti_data_bacteria$Var2),
                               grep("E", bnti_data_bacteria$Var1), 
                               grep("E", bnti_data_bacteria$Var2)),]

rcbc_data_palsa = rcbc_data_bac[-c(grep("[a-z]_S_", rcbc_data_bac$Var1), 
                               grep("[a-z]_S_", rcbc_data_bac$Var2),
                               grep("E", rcbc_data_bac$Var1),
                               grep("E", rcbc_data_bac$Var2)),]


# Determining ecological processes fractionation
eco.proc_palsa = data.frame(hom.sel = NA, 
                            hom.dis = NA, 
                            und = NA, 
                            dis.lim = NA, 
                            var.sel = NA,
                            row.names = "Palsa")

eco.proc_palsa$hom.sel = length(which(bnti_data_palsa$value <= -2)) / 
  length(bnti_data_palsa$value)

eco.proc_palsa$hom.dis = length(which(rcbc_data_palsa$value <= -0.95)) / 
  length(bnti_data_palsa$value)

eco.proc_palsa$und = length(which(abs(rcbc_data_palsa$value) < 0.95)) / 
  length(bnti_data_palsa$value)

eco.proc_palsa$dis.lim = length(which(rcbc_data_palsa$value >= 0.95)) / 
  length(bnti_data_palsa$value)

eco.proc_palsa$var.sel = length(which(bnti_data_palsa$value >= 2)) / 
  length(bnti_data_palsa$value)

eco.proc_palsa = melt(as.matrix(eco.proc_palsa))

eco.proc_palsa$value = (eco.proc_palsa$value*100)
```

##### Joining data
```{r}
microbe_data <- rbind(eco.proc_palsa, eco.proc_2, eco.proc_fen) %>% 
  mutate(Var2 = case_when(Var2 == 'hom.sel' ~ 'Homogeneous selection',
                          Var2 == 'hom.dis' ~ 'Homogenizing dispersal',
                          Var2 == 'und' ~ 'Undominated',
                          Var2 == 'dis.lim' ~ 'Dispersal limitation',
                          Var2 == 'var.sel' ~ 'Variable selection')) %>% 
  rename(Habitat = Var1,
         Assembly_process = Var2) %>% 
  mutate(Compound = 'Microbiome')

```

### Plotting

```{r}
all_eco_process <- rbind(data_metabolites, microbe_data) %>% 
  mutate(Compound = factor(Compound, 
                           levels = rev(c('Microbiome',
                                          'Bulk metabolites'))))

eco_process_bar <- ggplot(all_eco_process) +
  geom_col(aes(x = value,
               y = Compound,
               fill = Assembly_process),
           color = 'black') +
  scale_fill_manual(values = eco_process_colors) +
  theme_bw() +
  facet_grid(cols = vars(Habitat)) +
  guides(fill = guide_legend(nrow=2,byrow=TRUE)) +
  theme(strip.text.x = element_text(angle = 0, face = 'bold',
                                    color = 'white'),
        legend.position = 'bottom',
        axis.title.y = element_blank())

eco_process_bar

fills <- list_colors[c('Palsa', 'Bog', 'Fen')]

g <- ggplot_gtable(ggplot_build(eco_process_bar))
stripr <- which(grepl('strip-t', g$layout$name))

k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}

dev.off()
grid.draw(g)

eco_proc_final <- g
```

## Mantel Heatmap

### Loading biogeochemical data

```{r}
peat_envs <- read_csv("input/peat_geodata_5-20-23.csv")
```

```{r}

geo <- peat_envs %>% 
  select(SampleID, Month, Habitat, Depth, DepthAvg__, `T, soil (C)`, CtoN_wt, Precipitation_3day_accumulation_mm) %>% 
    rename(T_soil = `T, soil (C)`) %>% 
  rename_with(~str_remove(.x, '__')) 

```

### Load in bNTI
```{r}

weig <- read_csv(
  "input/Abisko_TWCD_bNTI_999_so_5_2_21.csv"
) %>% 
  column_to_rownames(var = "...1")

weig[upper.tri(weig)] = t(weig)[upper.tri(weig)]

```


```{r}
## Bacteria BNTI

samples_names <- read_csv(
  "input/metadata_matched_MAGs_data_11-29-21.csv"
)

bacteria_bnti <- read_csv("input/weighted_BNTI_bacteria.csv") %>% 
  column_to_rownames("...1")

bacteria_bnti <- bacteria_bnti[colnames(bacteria_bnti),] %>% 
  rename(!!(set_names(samples_names$names_bnti, samples_names$SampleID))) 

rownames(bacteria_bnti) <- colnames(bacteria_bnti)

# Reflecting  matrices

bacteria_r <- bacteria_bnti
bacteria_r[upper.tri(bacteria_bnti)] <-  t(bacteria_bnti)[upper.tri(bacteria_bnti)]
```

### Analysis by Habitat
#### Bog - ALL 

```{r}
#Filtering BOG data

geo_bog <- geo %>% 
  filter(str_detect(SampleID, "[a-z]_S"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm', 'bacteria')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_bog$SampleID, 
                                     colnames(bacteria_r) %in% geo_bog$SampleID,]
  } else {
    mat <- distance_func(weig, geo_bog, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth', 'C:N_ratio',
                        'precipitation', 'microbial_bNTI')


```

```{r}

mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(weig, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

mantel_stats_bog_all <- mantel_stats %>% 
  mutate(Type = 'Bulk')

```

#### FEN - ALL 

```{r}
#Filtering Fen data 
geo_Fen <- geo %>% 
  filter(str_detect(SampleID, "E"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm', 'bacteria')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_Fen$SampleID, 
                                     colnames(bacteria_r) %in% geo_Fen$SampleID,]
  } else {
    mat <- distance_func(weig, geo_Fen, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth', 'C:N_ratio',
                        'precipitation', 'microbial_bNTI')

```

```{r}
mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(weig, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

mantel_stats_fen_all <- mantel_stats %>% 
  mutate(Type = 'Bulk')

```

#### Palsa - ALL 

```{r}
#Filtering Palsa data 
geo_Palsa <- geo %>% 
  filter(str_detect(SampleID, "P"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm', 'bacteria')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_Palsa$SampleID, 
                                     colnames(bacteria_r) %in% geo_Palsa$SampleID,]
  } else {
    mat <- distance_func(weig, geo_Palsa, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth', 'C:N_ratio',
                        'precipitation', 'microbial_bNTI')

```

```{r}
mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(weig, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

mantel_stats_Palsa_all <- mantel_stats %>% 
  mutate(Type = 'Bulk')

```


#### Bacteria - Palsa

```{r}
#Filtering Palsa data 
geo_Palsa <- geo %>% 
  filter(str_detect(SampleID, "P"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_Palsa$SampleID, 
                                     colnames(bacteria_r) %in% geo_Palsa$SampleID,]
  } else {
    mat <- distance_func(bacteria_r, geo_Palsa, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth', 'C:N_ratio',
                        'precipitation')

```

```{r}
mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(bacteria_r, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

data_plot_Palsa_bacteria <- mantel_stats %>% 
  mutate(Habitat = 'Palsa',
         Type = 'Microbial \u03B2NTI')
```

#### Bacteria - Bog

```{r}
#Filtering Bog
geo_bog <- geo %>% 
  filter(str_detect(SampleID, "[a-z]_S"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_$SampbogleID, 
                                     colnames(bacteria_r) %in% geo_bog$SampleID,]
  } else {
    mat <- distance_func(bacteria_r, geo_bog, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth','C:N_ratio',
                        'precipitation')
```

```{r}
mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(bacteria_r, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

data_plot_Bog_bacteria <- mantel_stats %>% 
  mutate(Habitat = 'Bog',
         Type = 'Microbial \u03B2NTI')
```

#### Bacteria - Fen

```{r}
#Filtering fen data 
geo_Fen <- geo %>% 
  filter(str_detect(SampleID, "E"))

sel_columns <- c('T_soil', 'DepthAvg',
             'CtoN_wt', 'Precipitation_3day_accumulation_mm')

matrix_list <- map(sel_columns, function(x){
  if(x == 'bacteria'){
    mat <- bacteria_r[rownames(bacteria_r) %in% geo_Fen$SampleID, 
                                     colnames(bacteria_r) %in% geo_Fen$SampleID,]
  } else {
    mat <- distance_func(bacteria_r, geo_Fen, x)
  }
  
  return(mat)
})

names(matrix_list) <- c('soil_t', 'depth', 'C:N_ratio',
                        'precipitation')
```

```{r}
mantel_stats <- tibble(Geochemistry = character(), Mantel_stat = numeric(), p_value = numeric(), pval.adj = numeric(), .rows = 0)

for(i in 1:length(matrix_list)){
  
  mat <- as.data.frame(matrix_list[i])
  colnames(mat) <- str_remove(colnames(mat), '.*\\.')
  
  mantel_result <- mantel_fun(bacteria_r, mat)
  temp <- tibble(Geochemistry = names(matrix_list)[i], Mantel_stat = mantel_result$statistic, 
                 p_value = mantel_result$signif)
  
  mantel_stats <- rbind(mantel_stats, temp)
}

mantel_stats$pval.adj <- p.adjust(mantel_stats$p_value, method = 'bonferroni')

data_plot_Fen_bacteria <- mantel_stats %>% 
  mutate(Habitat = 'Fen',
         Type = 'Microbial \u03B2NTI')
```

### Joining and plotting

```{r}

mantel_stats_Palsa_all <- mantel_stats_Palsa_all %>% 
  mutate( Habitat = "Palsa") %>% 
  mutate(Type = 'Metabolite \u03B2NTI')

mantel_stats_bog_all <- mantel_stats_bog_all %>% 
  mutate( Habitat = "Bog") %>% 
  mutate(Type = 'Metabolite \u03B2NTI')

mantel_stats_fen_all <- mantel_stats_fen_all %>% 
  mutate( Habitat = "Fen") %>% 
  mutate(Type = 'Metabolite \u03B2NTI')


data_plot <- rbind(mantel_stats_Palsa_all, mantel_stats_bog_all, mantel_stats_fen_all,
                   data_plot_Palsa_bacteria, data_plot_Bog_bacteria,
                   data_plot_Fen_bacteria) %>% 
  mutate(Signif = case_when(pval.adj <= 0.0001 ~ '****',
                            pval.adj <= 0.001 ~  '***',
                            pval.adj <= 0.01 ~    '**',
                            pval.adj <= 0.05 ~ '*')) %>% 
  mutate(Habitat = factor(Habitat, levels = c('Palsa', 'Bog', 'Fen'))) %>% 
  rbind(list('microbial_bNTI', NA, NA, NA, 'Microbial \u03B2NTI', 'Palsa', NA)) %>% 
  rbind(list('microbial_bNTI', NA, NA, NA, 'Microbial \u03B2NTI', 'Bog', NA)) %>% 
  rbind(list('microbial_bNTI', NA, NA, NA, 'Microbial \u03B2NTI', 'Fen', NA))


mantel_heatmap <- data_plot %>% 
  mutate(Geochemistry = case_when(Geochemistry == 'soil_t' ~ 'Soil temperature',
                                  Geochemistry == 'depth' ~ 'Depth',
                                  Geochemistry == 'microbial_bNTI' ~ 'Microbial \u03B2NTI',
                                  Geochemistry == 'C:N_ratio' ~ 'C:N ratio',
                                  Geochemistry == 'precipitation' ~ 'Precipitation',
                                  TRUE ~ Geochemistry),
         Geochemistry = factor(Geochemistry, levels = c('Depth',
                                                        'Soil temperature',
                                                        'C:N ratio',
                                                        'Precipitation',
                                                        'Microbial \u03B2NTI'))) %>%  
  ggplot(aes(x = Type, y = Geochemistry, fill = Mantel_stat))+
  geom_tile(color = "white")+
  scale_fill_distiller(palette = 'RdBu', 
                       limits = c(-0.55, 0.55), na.value = 'gray80')+
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  geom_text(aes(label = Signif))+
  labs(fill = 'Mantel \nStatistic')+
  theme(plot.title = element_text(face="bold", hjust = 0.5),
        strip.text = element_text(face = "bold", color = "white"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_blank()) +
  facet_wrap(~Habitat)

mantel_heatmap

# Changing facet strips colors

fills <- list_colors[c('Palsa', 'Bog', 'Fen')]

g <- ggplot_gtable(ggplot_build(mantel_heatmap))
stripr <- which(grepl('strip-t', g$layout$name))

k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}

dev.off()
grid.draw(g)

t2 <- as_ggplot(g)

```

## Figure 1 -Join

```{r}

partial_1 <- ggarrange(bp_group_vio_bacteria, bp_group_vio_so,
                       labels = "AUTO")

partial_1

partial_2 <- ggarrange(depth_den, month_den, common.legend = TRUE, legend = "bottom",
                       labels = c("C", "D"))


partial_2


partial_3 <- ggarrange(partial_1, partial_2,
                       ncol = 1)

partial_3


partial_4 <- ggarrange(eco_proc_final, t2,
                       ncol = 1,
                       labels = c("E", "F"))

partial_4


final_fig1 <- ggarrange(partial_3, partial_4)

final_fig1


ggsave('output/Figure1_final.png', final_fig1, dpi = 300, height = 8, width = 15)


```







