---
title: "Metabolome paper - Figure 2"
author: "Viviana Freire Zapata"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Loading libraries

```{r message=FALSE, warning=FALSE}
library(igraph)
library(patchwork)
library(ggnewscale)
library(rstatix)
library(ggpubr)
library(ggraph)
library(colourvalues)
library(tidygraph)
library(tidyverse)
```

# Load data

## Microbial correlation

```{r}
habitat <- c('Bog', 'Fen', 'Palsa')
micro_cor_files <- list.files(path = '../Metabolite_clusters/output/',
                              pattern = 'sig_corr_taxa',
                              recursive = TRUE,
                              full.names = TRUE)

names(micro_cor_files) <- habitat

micro_cor <- map(micro_cor_files, function(x){
  df <- read_csv(x) %>% 
    rename(to = Microbe,
           pval_adj = p_adj)
})
```

## GHG correlations

```{r}
gas_cor_files <- list.files(path = '../Metabolite_clusters/output/',
                            pattern = 'gas_correlations',
                            recursive = TRUE,
                            full.names = TRUE)

names(gas_cor_files) <- habitat

gas_cor <- map(gas_cor_files, function(x){
  df <- read_csv(x) %>% 
    rename(to = env_variable) 
})
```

## Environmental correlations

```{r}
env_cor_files <- list.files(path = '../Metabolite_clusters/output/',
                            pattern = 'env_correlations',
                            recursive = TRUE,
                            full.names = TRUE)

names(env_cor_files) <- habitat

env_cor <- map(env_cor_files, function(x){
  df <- read_csv(x) %>% 
    rename(to = env_variable)
})
```

## BNTI feature data (with taxonomy)

```{r}
micro_bnti_files <- list.files(path = '../Assembly_feature_level_microbiome/output/',
                               pattern = 'feature_microbe_tax',
                               recursive = TRUE,
                               full.names = TRUE)
names(micro_bnti_files) <- habitat

micro_bnti <- map(micro_bnti_files, function(x){
  df <- read_csv(x) %>% 
    filter(Direction != 'Insignificant')
})

micro_bnti_for_filt <- map(micro_bnti_files, function(x){
  df <- read_csv(x)
})
```

## Representative features BNTI data

```{r}
bnti_counts_files <- list.files(path = '../Metabolite_clusters/output/',
                                pattern = 'rep_features',
                                recursive = TRUE,
                                full.names = TRUE)

bnti_counts_files <- bnti_counts_files[str_detect(bnti_counts_files, 'bnti', negate = TRUE)]

names(bnti_counts_files) <- habitat

bnti_counts <- imap(bnti_counts_files, function(x, y){
  df <- read_csv(x) %>% 
    group_by(cluster, Direction) %>% 
    count() %>% 
    group_by(cluster) %>% 
    mutate(perc = n/sum(n) * 100,
           Habitat = y)
})
```

# Preprocessing

## Filtering bacterial correlations

```{r}
micro_cor_filt <- map2(micro_cor, micro_bnti_for_filt, function(x, y){
  df <- x %>% 
    filter(to %in% y$Member)
})

micro_cor_sel <- map(micro_cor_filt, function(df){
  df %>% 
    select(cluster, to, rho, pvalue, pval_adj)
})
```

# Plots colors

```{r}
taxa <- do.call(rbind, micro_cor_filt) %>% 
  select(to, Phylum) %>% 
  distinct() 

taxa_list <- taxa %>% 
  pull(Phylum) %>% 
  unique() %>% 
  str_remove(., 'p__')

phyl_colors <- set_names(c(ggpubr::get_palette('jco', length(sort(taxa_list))), 'transparent'),
                         nm = c(sort(taxa_list), ''))

habitat_colors <- c('Palsa' = '#703C1B', 'Bog' = '#058000', 'Fen' = '#0001FF')

contrib_colors <- set_names(ggpubr::get_palette(palette = 'RdYlBu', 5),
                            nm = c('Sig. High', 'High', 'Insignificant', 'Low', 'Sig. Low'))
```

# Correlation network

## Building network with igraph

```{r}
network_graph <- pmap(
  list(micro_cor_sel, gas_cor, env_cor, bnti_counts, micro_bnti),
  function(micro, gas, env, bnti, m_bnti){
    
    edges <- rbind(micro, gas, env) %>% 
      filter(pval_adj < 0.05) %>% 
      rename(from = cluster) %>% 
      mutate(arho = abs(rho),
             width = .5 + ((arho - min(arho))*(4-.5))/(max(arho) - min(arho)),
             color = color_values(rho, palette = 'blue2red'))
    
    nodes <- tibble(Name = c(unique(edges$to),
                             unique(edges$from))) %>% 
      left_join(taxa, by = c('Name' = 'to')) %>% 
      mutate(type = case_when(Name %in% edges$from ~ 'Met. Cluster',
                              Name %in% micro$to ~ 'Bacteria',
                              Name %in% gas$to ~ 'Env. Factor',
                              Name %in% env$to ~ 'Env. Factor'),
             type = factor(type, levels = c('Met. Cluster', 'Bacteria', 'Env. Factor')),
             # shape = case_when(type == 'cluster' ~ 'pie',
             #                   Name %in% m_bnti$Member ~ 'star',
             #                   type == 'micro' ~ 'circle',
             #                   TRUE ~ 'square'),
             Phylum = case_when(type == 'Bacteria' ~ str_remove(Phylum, 'p__')),
             Phylum = ifelse(is.na(Phylum), '', Phylum),
             Phylum = factor(Phylum, levels = str_remove(names(phyl_colors), 'p__')),
             Other_nodes = case_when(type == 'Met. Cluster' ~ 'Met. Cluster',
                                     type == 'Env. Factor' ~ 'Env. Factor'),
             Other_nodes = factor(Other_nodes, levels = c('Met. Cluster', 'Env. Factor')),
             size = case_when(type == 'Met. Cluster' ~ 6,
                              type == 'Bacteria' ~ 4,
                              TRUE ~ 6),
             label = ifelse(type == 'Bacteria', NA, Name),
             label = str_replace(label, 'Precipitation_3day_accumulation_mm', 'Precipitation'),
             label = str_replace(label, 'DepthAvg__', 'Depth'),
             label = str_replace(label, 'T, soil \\(C\\)', 'T. soil'))
    
    g <- graph_from_data_frame(edges, vertices = nodes, directed = FALSE)
    nodes <- nodes %>% 
      rename(id = Name)
    
    return(list(nodes = nodes,
                edges = edges,
                graph = g))
  }
)
```

## Plotting network with gg_graph

```{r}

legend_p <- tibble(Taxa = names(phyl_colors),
                   `Node Type` = c(rep('Met. Cluster', 4), 
                                   rep('Bacteria', 4),
                                   rep('Env. Factor', 3)),
                   rho = seq(-0.7, 0.7, length.out = 11)) %>% 
  mutate(Taxa = factor(Taxa, levels = names(phyl_colors)),
         `Node Type` = factor(`Node Type`, 
                              levels = c('Met. Cluster', 'Bacteria', 'Env. Factor'))) %>% 
  ggplot() +
  geom_point(aes(x = Taxa,
                 y = 1,
                 color = Taxa,
                 shape = `Node Type`),
             alpha = 0.75) +
  geom_boxplot(aes(x = Taxa,
                   y = 1,
                   fill = rho)) +
  scale_color_manual(values = phyl_colors) +
  scale_shape_manual(values = c('Met. Cluster' = 15, 
                                'Bacteria' = 16, 
                                'Env. Factor' = 17)) +
  scale_fill_distiller(palette = 'RdBu', limits = c(-0.7, 0.7)) +
  guides(color = guide_legend(nrow = 2, direction = 'vertical'),
         shape = guide_legend(nrow = 2, direction = 'vertical'),
         fill = guide_colorbar(direction = 'horizontal', title.position = 'top')) +
  labs(fill = "Spearman's\nrho") +
  theme_bw() +
  theme(legend.position = 'bottom',
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 7),
        legend.key.size = unit(.75, 'line'),
        legend.spacing.x = unit(0, 'line'),
        legend.spacing.y = unit(0, 'line'))

legend_p

legend <- get_legend(legend_p)

set.seed(123)
gg_network <- imap(network_graph, function(x, y){
  gg <- tbl_graph(nodes = x$nodes, edges = x$edges, node_key = 'id')
  
  ggraph(gg, layout = 'fr') +
    geom_edge_arc(strength = .1,
                  aes(edge_color = rho,
                      edge_width = arho)) +
    scale_edge_color_distiller(palette = 'RdBu', limits = c(-.7, .7)) +
    scale_edge_width(range = c(0.3, .7)) +
    guides(edge_width = 'none') +
    geom_node_point(aes(color = Phylum,
                        shape = type),
                    size = 2) +
    scale_color_manual(values = phyl_colors, drop = FALSE) +
    #guides(color = guide_legend(override.aes = list(color = phyl_colors))) +
    scale_shape_manual(values = c('Met. Cluster' = 15, 
                                  'Bacteria' = 16, 
                                  'Env. Factor' = 17),
                       drop = FALSE) +
    new_scale_color() +
    geom_node_point(aes(color = Other_nodes,
                        shape = type),
                    size = 3, show.legend = FALSE) +
    guides(shape = guide_legend(override.aes = list(color = c('black', 
                                                              'black', 'forestgreen'),
                                                    fill = c('black', 
                                                             'black', 'forestgreen')))) +
    scale_color_manual(values = c('Met. Cluster' = 'black', 
                                  'Env. Factor' = 'forestgreen'),
                       na.value = 'transparent', drop = FALSE) +
    geom_node_label(aes(label = label), hjust = 0.5, nudge_y = -0.5, 
                    repel = TRUE, size = 1.8) +
    labs(title = y,
         shape = 'Node\nType',
         edge_color = "Spearman's\nrho") +
    theme_void() +
    theme(legend.position = 'none',
          panel.border = element_rect(color = 'black', fill = NA),
          plot.title = element_text(face = 'bold', hjust = 0.5, size = 8))
})

gg_network_final <- gg_network$Palsa + gg_network$Bog + gg_network$Fen 


```

# Biochemical index plot

## Data wrangling

```{r}
rep_df <- imap(bnti_counts_files, function(x, y){
  df <- read_csv(x) %>% 
    mutate(Habitat = y)
})

rep_df_final <- reduce(rep_df, rbind) %>% 
  mutate(cluster = paste0('Cluster', cluster),
         Habitat = factor(Habitat, levels = c('Palsa', 'Bog', 'Fen')))


idx_stat <- rep_df_final %>% 
  select(Habitat, cluster, NOSC, AI_Mod, DBE_O) %>% 
  pivot_longer(!c(Habitat, cluster), names_to = 'index', values_to = 'value') %>%
  group_by(Habitat, index) %>% 
  wilcox_test(value ~ cluster, p.adjust.method = "bonferroni") %>% 
  add_xy_position(scales = 'free', step.increase = 1) %>% 
  mutate(p.adj = paste0("P = ", p.adj))



```

## Plotting

```{r}
idx_plot <- rep_df_final %>% 
  select(Habitat, cluster, NOSC, AI_Mod, DBE_O) %>% 
  pivot_longer(!c(Habitat, cluster), names_to = 'index', values_to = 'value') %>% 
  ggplot(aes(x = cluster,
             y = value,
             fill = Habitat)) +
  geom_violin(alpha = 0.5, linewidth = .1) +
  geom_boxplot(width = 0.1, linewidth = .1, outlier.size = .2) +
  facet_grid(cols = vars(Habitat),
             rows = vars(index),
             scales = 'free',
             space = 'free_x') +
  scale_fill_manual(values = habitat_colors) +
  stat_pvalue_manual(data = idx_stat, label = 'p.adj', hide.ns = TRUE,
                     size = 1.5) +
  scale_y_continuous(expand = expansion(mult = 0.1)) +
  theme_bw() +
  theme(strip.text.y = element_text(size = 6),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.y = element_text(size = 7),
        strip.text.x = element_text(size = 6),
        legend.position = 'none')

idx_plot

```

# Feature contribution barplot

```{r}
bnti_counts_df <- reduce(bnti_counts, rbind)

contrib_plot <- bnti_counts_df %>% 
  mutate(cluster = paste0('Cluster', cluster),
         Direction = factor(Direction, 
                            levels = c('Sig. High', 'High', 'Insignificant', 'Low', 'Sig. Low')),
         Habitat = factor(Habitat, levels = c('Palsa', 'Bog', 'Fen'))) %>% 
  ggplot() +
  geom_col(aes(x = cluster,
               y = perc,
               fill = Direction),
           color = 'black',
           linewidth = .1) +
  facet_grid(cols = vars(Habitat),
             scales = 'free',
             space = 'free_x') +
  labs(fill = 'Feature Contribution',
       y = 'Percentage [%]') +
  scale_y_continuous(labels = ~ scales::percent(.x, scale = 1)) +
  scale_fill_manual(values = contrib_colors) +
  guides(fill = guide_legend(title.position = 'top')) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.y = element_text(size = 7),
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 7),
        strip.text = element_text(size = 6),
        legend.position = 'bottom',
        legend.key.size = unit(.5, 'line'),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, 0))

contrib_plot
```

# Elemental composition barplots

```{r}
class_plot <- rep_df_final %>% 
  group_by(cluster, Habitat) %>% 
  count(El_comp) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = cluster,
               y = perc,
               fill = El_comp),
           color = 'black',
           linewidth = .1) +
  facet_grid(cols = vars(Habitat),
             scales = 'free',
             space = 'free_x') +
  labs(fill = 'Elemental Composition',
       y = 'Percentage [%]') +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = ggpubr::get_palette(palette = 'Set3', 10)) +
  guides(fill = guide_legend(title.position = 'top')) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        axis.title.y = element_text(size = 7),
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 7),
        strip.text = element_text(size = 6), 
        legend.position = 'bottom',
        legend.key.size = unit(.5, 'line'),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, 0))


class_plot
```

# Finalizing Figure 2

```{r}

des <- "
AABBCC
DDDDDD
EEEFFF
EEEGGG
"

figure_2 <-  gg_network_final + legend + free(idx_plot) + 
  free(contrib_plot) + free(class_plot) +
  plot_layout(heights = c(5, 1.5, 3, 3), design = des) +
  plot_annotation(tag_levels = list(c('A', '', '', '', 'B', 'C', 'D'))) &
  theme(plot.tag = element_text(size = 7))


figure_2

ggsave('output_updated/Figure_2.png', figure_2, dpi = 300,  width = 180, 
       height = 186, units = 'mm')

ggsave('output_updated/Figure_2.eps', figure_2, 
       device = cairo_ps, fallback_resolution = 300,
       width = 180, height = 186, units = 'mm')
```

















