---
title: "Metabolome paper - Supplementary Figure 2"
author: "Viviana Freire Zapata"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Loading libraries

```{r message=FALSE, warning=FALSE}
library(ggpubr)
library(patchwork)
library(readxl)
library(ggtext)
library(ggh4x)
library(patchwork)
library(tidyverse)
```

# Load data

All data is only for the **Bog Habitat**

## Metadata

```{r}
metadata_bog <- read_csv("../Assembly_feature_level_microbiome/input/metadata_matched_MAGs_data.csv") %>% 
  filter(Habitat == 'Bog') %>% 
  select(c(SampleID, names_bnti)) %>% 
  rename(SampleID_1 = names_bnti) 
```


## OTU table

```{r}

otu_table_bog <- read_csv("../Assembly_feature_level_microbiome/input/otu_normalized.csv")  %>% 
  column_to_rownames(var = "genome") %>% 
  select(all_of(metadata_bog$SampleID_1)) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "SampleID_1") %>% 
  left_join(metadata_bog, by = "SampleID_1") %>% 
  select(SampleID, everything(), -SampleID_1)
```


## Microbes significantly correlated with BNTI clusters

```{r}
bog_sig_cor_taxa <- read_csv('../Metabolite_clusters/output/bog_sig_corr_taxa.csv')
```

## BNTI_feature contribution 

```{r}
bog_bnti_contrib <- read_csv('../Assembly_feature_level_microbiome/output/bog_feature_microbe_tax_3-22-22.csv') %>% 
  select(Member, Value, Direction)

```

## CO2 data

```{r}
bog_gas <- read_csv('../input/bog_matrix_3-1-22.csv') %>% 
  select(SampleID, Month, Habitat, CH4.mM__, CO2.mM__) %>% 
  drop_na()
```

## DRAM Annotation

```{r}
bog_annotation_carbon <- read_xlsx('../input/1864_97percentmags_metabolism_summary.xlsx', 
                                   sheet = 'carbon utilization') %>% 
  mutate(group = 'Carbon Utilization')

bog_annotation_energy <- read_xlsx('../input/1864_97percentmags_metabolism_summary.xlsx',
                                   sheet = 'Energy') %>% 
  mutate(group = 'Energy')

bog_annotation <- rbind(bog_annotation_carbon, bog_annotation_energy) %>% 
  pivot_longer(!c(gene_id, gene_description, module, header, subheader, group),
               names_to = 'genome', values_to = 'value') %>% 
  mutate(Presence = if_else(value > 0, 'Present', 'Absent'))


```


# Plots colors

```{r}
contrib_colors <- set_names(ggpubr::get_palette('RdYlBu', 5),
                            nm = c('Sig. High', 'High', 'Insignificant', 'Low', 'Sig. Low'))
```


# Data wrangling

## Getting correlated features with a significant contribution

```{r}
bog_sig_cor_contrib <- bog_sig_cor_taxa %>% 
  left_join(bog_bnti_contrib, by = c('Microbe' = 'Member')) %>% 
  filter(Direction != 'Insignificant') %>% 
  drop_na(Direction) %>% 
  select(Microbe, Direction, Phylum, Genus) %>% 
  distinct()
```

```{r}
processed_gas_table <- otu_table_bog %>% 
  filter(SampleID %in% bog_gas$SampleID) %>% 
  pivot_longer(!SampleID, names_to = 'genome', values_to = 'abundance') %>% 
  filter(abundance > 0) %>% 
  inner_join(bog_gas, by = 'SampleID') %>% 
  inner_join(bog_sig_cor_contrib, by = c('genome' = 'Microbe'))
```

# CO2

## Calculating correlation

```{r}
bog_co2_corr <- processed_gas_table %>% 
  group_by(genome) %>% 
  nest() %>% 
  mutate(co2_rho = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CO2.mM__, type = 'spearman')
    corr$r[1,2]
  })) %>% 
  mutate(co2_pval = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CO2.mM__, type = 'spearman')
    corr$P[1,2]
  })) %>% 
  mutate(pval_adj = p.adjust(co2_pval, method = 'fdr')) %>% 
  filter(pval_adj < 0.05)
```

## Plotting 

Only significant

```{r}
bog_co2_plot <- processed_gas_table %>% 
  filter(genome %in% bog_co2_corr$genome) %>% 
  mutate(Phylum = str_remove(Phylum, 'p__'),
         Genus = str_remove(Genus, 'g__'),
         strip_label = paste0(Phylum, ' ',
                              '*', Genus, '*', '<br>',
                              '(', genome, ')')) %>% 
  ggplot(aes(x = abundance,
             y = CO2.mM__,
             color = Direction)) +
  geom_point(size = 0.5) +
  geom_smooth(method = 'lm', se = FALSE, linewidth = .3) +
  labs(title = 'Correlations with CO<sub>2</sub>',
       x = 'Normalized abundance',
       y = 'CO<sub>2</sub> [mM]',
       color = 'Contribution') +
  scale_color_manual(values = contrib_colors) +
  stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = '~`,`~')), 
           method = "spearman",
           color = 'black',
           size = 2,
           show.legend = FALSE) +
  theme_bw() +
  theme(plot.title = element_markdown(face = 'bold', hjust = 0.5),
        axis.title.y = element_markdown(),
        strip.text.x = element_markdown()) +
  facet_wrap(~strip_label, scales = 'free_x', ncol = 3)

bog_co2_plot  
```


# CH4

## Calculating correlation

```{r}
bog_ch4_corr <- processed_gas_table %>% 
  group_by(genome) %>% 
  nest() %>% 
  mutate(co2_rho = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CH4.mM__, type = 'spearman')
    corr$r[1,2]
  })) %>% 
  mutate(co2_pval = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CH4.mM__, type = 'spearman')
    corr$P[1,2]
  })) %>% 
  mutate(pval_adj = p.adjust(co2_pval, method = 'fdr')) %>% 
  filter(pval_adj < 0.05)
```

## Plotting 

Only significant

```{r}
bog_ch4_plot <- processed_gas_table %>% 
  filter(genome %in% bog_ch4_corr$genome) %>% 
  mutate(Phylum = str_remove(Phylum, 'p__'),
         Genus = str_remove(Genus, 'g__'),
         strip_label = paste0(Phylum, ' ',
                              '*', Genus, '*', '<br>',
                              '(', genome, ')')) %>% 
  ggplot(aes(x = abundance,
             y = CH4.mM__,
             color = Direction)) +
  geom_point(show.legend = FALSE, size = 0.5) +
  geom_smooth(method = 'lm', se = FALSE, show.legend = FALSE, linewidth = .3) +
  labs(title = 'Correlations with CH<sub>4</sub>',
       y = 'CH<sub>4</sub> [mM]',
       x = 'Normalized abundance') +
  scale_color_manual(values = contrib_colors) +
  stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = '~`,`~')), 
           method = "spearman",
           size = 2,
           color = 'black',
           show.legend = FALSE) +
  theme_bw() +
  theme(plot.title = element_markdown(face = 'bold', hjust = 0.5),
        axis.title.y = element_markdown(),
        # legend.position = 'none',
        strip.text.x = element_markdown()) +
  facet_wrap(~strip_label, scales = 'free_x', ncol = 2)


bog_ch4_plot  
```

# Annotation plot

```{r}
bog_annotation_filt <- bog_annotation %>% 
  filter(genome %in% c(bog_co2_corr$genome, bog_ch4_corr$genome),
         header != 'CAZY',
         Presence == 'Present') %>% 
  mutate(
    inner_level = case_when(
      group == 'Carbon Utilization' & is.na(subheader) ~ header,
      group == 'Carbon Utilization' & !is.na(subheader) ~ subheader,
      group == 'Energy' ~ module
    ),
    outer_level = case_when(
      group == 'Carbon Utilization' & !is.na(header) ~ header,
      group == 'Carbon Utilization' & subheader == 'Polyphenolics Cleavage' ~ 'Polyphenols',
      group == 'Energy' ~ header
    )
    
  )

bog_annotation_plot_carbon <- bog_annotation_filt %>% 
  filter(group == 'Carbon Utilization') %>% 
  mutate(genome = factor(genome, levels = c('20110800_S2M_3',
                                            '20120800_E1D_14',
                                            '20120700_S1D_25',
                                            '20140700_E14_11',
                                            '20120600_S1S_4'))) %>% 
  ggplot() +
  geom_tile(aes(x = genome,
                y = gene_id),
            fill = 'tomato3',
            color = 'white') +
  theme_bw() +
  facet_grid2(inner_level ~ .,
               scales = 'free_y',
               space = 'free',
               strip = strip_nested(size = 'variable',
                                    background_y = list(
                                      element_rect(fill = 'white')
                                    ),
                                    by_layer_y = TRUE)) +
  labs(title = 'Carbon Utilization') +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y = element_text(angle = 0),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.placement = 'outside')  

bog_annotation_plot_carbon

bog_annotation_plot_energy <- bog_annotation_filt %>% 
  filter(group == 'Energy',
         outer_level != 'Nitrogen',
         !(outer_level == 'Sulfur' & is.na(inner_level))) %>% 
  mutate(genome = factor(genome, levels = c('20110800_S2M_3',
                                            '20120800_E1D_14',
                                            '20120700_S1D_25',
                                            '20140700_E14_11',
                                            '20120600_S1S_4'))) %>% 
  ggplot() +
  geom_tile(aes(x = genome,
                y = gene_id),
            fill = 'purple4',
            color = 'white') +
  theme_bw() +
  facet_nested(outer_level + inner_level ~ .,
               scales = 'free_y',
               space = 'free',
               nest_line = element_line(linetype = 2),
               strip = strip_nested(size = 'variable',
                                    background_y = list(
                                      element_rect(fill = 'white'),
                                      NULL
                                    ),
                                    by_layer_y = TRUE)) +
  labs(title = 'Energy') +
  # scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  theme(plot.title = element_text(face = 'bold', hjust = 0.5),
        axis.title = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y = element_text(angle = 0),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(.1, 'lines'),
        strip.placement = 'outside')  

bog_annotation_plot_energy
```

# Merging final figure

```{r}
sub1 <- bog_co2_plot + bog_ch4_plot +
  plot_layout(guides = 'collect',
              widths = c(2.75, 2)) +
  plot_annotation(tag_levels = 'A') &
  theme(text = element_text(size = 5),
        plot.tag = element_text(size = 8))

sub2 <- bog_annotation_plot_carbon + bog_annotation_plot_energy +
  plot_annotation(tag_levels = list('C', 'D')) &
  theme(text = element_text(size = 5),
        plot.tag = element_text(size = 8),
        axis.text.y = element_text(size = 3))

final <- ggarrange(sub1, sub2, nrow = 2, heights = c(1, 3))

ggsave('output/Figure_3.png', final, dpi = 300, height = 210, 
       width = 180, units = 'mm')
```


# GHG correlations without filtering MAGs that contributed to assembly


```{r}
processed_gas_table <- otu_table_bog %>%
  filter(SampleID %in% bog_gas$SampleID) %>%
    pivot_longer(!SampleID, names_to = 'genome', values_to = 'abundance') %>%
    filter(abundance > 0) %>%
    inner_join(bog_gas, by = 'SampleID') %>% 
  filter(genome %in% bog_sig_cor_taxa$Microbe)
```


# CO2 concentration

## Calculating correlation

```{r}
bog_co2_corr <- processed_gas_table %>% 
  group_by(genome) %>% 
  nest() %>% 
  mutate(co2_rho = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CO2.mM__, type = 'spearman')
    corr$r[1,2]
  })) %>% 
  mutate(co2_pval = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CO2.mM__, type = 'spearman')
    corr$P[1,2]
  })) %>% 
  mutate(pval_adj = p.adjust(co2_pval, method = 'fdr')) %>% 
  filter(pval_adj < 0.05)

```

## Joining with Taxonomy

```{r}

otu_table_bog <- read_csv("../Assembly_feature_level_microbiome/input/otu_normalized.csv") 

otu_taxa <- otu_table_bog %>% 
  dplyr::select(c(genome, GTDB_v2_r207)) %>% 
  separate_wider_delim(GTDB_v2_r207, delim = ";", names = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

bog_co2_corr_tax <- bog_co2_corr %>% 
  left_join(otu_taxa, by = "genome") %>% 
  left_join(bog_bnti_contrib, by = c("genome" = "Member"))

temp <- bog_co2_corr_tax %>% 
  select(-data) %>% 
  mutate(co2_rho = as.numeric(co2_rho),
         co2_pval = as.numeric(co2_pval))

write_csv(temp, "tables/Bog_mags_corr_CO2.csv")
```

# CH4

## Calculating correlation

```{r}
bog_ch4_corr <- processed_gas_table %>% 
  group_by(genome) %>% 
  nest() %>% 
  mutate(ch4_rho = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CH4.mM__, type = 'spearman')
    corr$r[1,2]
  })) %>% 
  mutate(ch4_pval = map(data, function(df){
    corr <- Hmisc::rcorr(df$abundance, df$CH4.mM__, type = 'spearman')
    corr$P[1,2]
  })) %>% 
  mutate(pval_adj = p.adjust(ch4_pval, method = 'fdr')) %>% 
  filter(pval_adj < 0.05)
```

## Joining with Taxonomy

```{r}

bog_ch4_corr_tax <- bog_ch4_corr %>% 
  left_join(otu_taxa, by = "genome")%>% 
  left_join(bog_bnti_contrib, by = c("genome" = "Member"))


temp <- bog_ch4_corr_tax %>% 
  dplyr::select(!data) %>% 
  mutate(ch4_rho = as.numeric(ch4_rho),
         ch4_pval = as.numeric(ch4_pval))


write_csv(temp, "tables/Bog_mags_corr_CH4.csv")
```


## Plotting 

Only significant

```{r}

bog_ch4_plot <- processed_gas_table %>% 
  left_join(otu_taxa, by = "genome") %>% 
  filter(genome %in% bog_ch4_corr_tax$genome) %>% 
  mutate(Phylum = str_remove(Phylum, 'p__'),
         Genus = str_remove(Genus, 'g__'),
         strip_label = paste0(Phylum, '<br>',
                              '*', Genus, '*', '<br>',
                              '(', genome, ')')) %>% 
  ggplot(aes(x = abundance,
             y = CH4.mM__,
             #color = Direction
             )) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  labs(title = 'Correlations with CH<sub>4</sub>',
       y = 'CH<sub>4</sub> [mM]',
       x = 'Normalized abundance') +
  #scale_color_manual(values = contrib_colors) +
  stat_cor(aes(label = paste(after_stat(rr.label), after_stat(p.label), sep = '~`,`~')), 
           method = "spearman",
           color = 'black',
           show.legend = FALSE) +
  theme_bw() +
  theme(plot.title = element_markdown(face = 'bold', hjust = 0.5),
        axis.title.y = element_markdown(),
        legend.position = 'none',
        strip.text.x = element_markdown()) +
  facet_wrap(~strip_label, scales = 'free_x')


bog_ch4_plot  


ggsave("output/Bog_cluster_correlations_CH4_3_6_2024.png", dpi = 300, height = 13.5, width = 15)


```






