---
title: "Metabolome paper - BNTI-feature clusters"
author: "Viviana Freire"
date: "`r Sys.Date()`"
output: html_document
---

# Loading libraries

```{r message=FALSE, warning=FALSE}
library(cluster)
library(party)
library(tidyverse)
```

# Functions

```{r}
test_k <- function(dist_matrix, k_init, k_final){
  # This functions test several values of K to find which maximizes ASV
  for(k in k_init:k_final){
    clustering <- pam(dist_matrix, k = k)
    sil <-  silhouette(clustering)
    avs <- mean(sil[,3])
    print(paste0('k = ', k, '; AVS = ', avs))
    
  }
}

cluster_rep_features <- function(dist_matrix, k){
  # Function to calculate the final clusters and get the representative features of each
  
  final_clusters <- pam(dist_matrix, k = k)
  
  cluster_df <- tibble(featureid = as.double(names(final_clusters$clustering)),
                       cluster = final_clusters$clustering) %>% 
    mutate(cluster = factor(cluster))
  
  # Getting representatives
  sil_data <- silhouette(final_clusters)
  
  sil_df <- tibble(featureid = as.double(rownames(sil_data)),
                   cluster = sil_data[,1],
                   sil_width = sil_data[,3]) %>% 
    mutate(cluster = factor(cluster))
  
  rep_features <- sil_df %>% 
    group_by(cluster) %>% 
    filter(sil_width > mean(sil_width))
  
  res <- list(all_features = cluster_df,
              representatives = rep_features)
}
```

# Clustering data

## Palsa

### Load datasets

```{r}
palsa_mt <- read_csv("input/bNTI_feat_TWCD_within_palsa_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1')

# Using only masses present in more than 50% of samples
palsa_mt_filt <- palsa_mt[rowSums(!is.na(palsa_mt)) > 15,]

palsa_chem <- read_csv('input/Palsa_feature_TWCD_mol_8-2-22.csv')
```

### Creating dissimilarity matrix

```{r}
palsa_dist <- daisy(palsa_mt_filt, metric = 'manhattan')
```


### Finding best k value

```{r cache=TRUE}
test_k(dist_matrix = palsa_dist, k_init = 3, k_final = 10) 
```

Best value for k is **3**

### Getting representative features

```{r}
k <- 3

palsa_clusters <- cluster_rep_features(dist_matrix = palsa_dist, k = k)

rep_features_palsa <- palsa_clusters$representatives %>% 
  left_join(palsa_chem, by = c('featureid' = 'Member')) %>% 
  ungroup()

write_csv(rep_features_palsa, 'output/rep_features_palsa.csv')

rep_features_palsa_bnti <- palsa_mt_filt %>% 
  rownames_to_column(var = 'featureid') %>% 
  mutate(featureid = as.double(featureid)) %>% 
  right_join(rep_features_palsa, by = 'featureid') %>% 
  select(featureid, cluster, all_of(colnames(palsa_mt_filt)))

write_csv(rep_features_palsa_bnti, 'output/rep_features_palsa_bnti.csv')
```

## Bog

### Load datasets

```{r}
bog_mt <- read_csv("input/bNTI_feat_TWCD_within_bog_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1')

# Using only masses present in more than 50% of samples
bog_mt_filt <- bog_mt[rowSums(!is.na(bog_mt)) > 14,]

bog_chem <- read_csv('input/bog_feature_TWCD_mol_8-2-22.csv')
```

### Creating dissimilarity matrix

```{r}
bog_dist <- daisy(bog_mt_filt, metric = 'manhattan')
```

### Finding best k value

```{r cache=TRUE}
test_k(dist_matrix = bog_dist, k_init = 3, k_final = 10) 
```

Best value for k is **4**

### Getting representative features

```{r}
k <- 4

bog_clusters <- cluster_rep_features(dist_matrix = bog_dist, k = k)

rep_features_bog <- bog_clusters$representatives %>% 
  left_join(bog_chem, by = c('featureid' = 'Member')) %>% 
  ungroup()

write_csv(rep_features_bog, 'output/rep_features_bog.csv')

rep_features_bog_bnti <- bog_mt_filt %>% 
  rownames_to_column(var = 'featureid') %>% 
  mutate(featureid = as.double(featureid)) %>% 
  right_join(rep_features_bog, by = 'featureid') %>% 
  select(featureid, cluster, all_of(colnames(bog_mt_filt)))

write_csv(rep_features_bog_bnti, 'output/rep_features_bog_bnti.csv')
```

## Fen

### Load datasets

```{r}
fen_mt <- read_csv("input/bNTI_feat_TWCD_within_fen_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1')

# Using only masses present in more than 50% of samples
fen_mt_filt <- fen_mt[rowSums(!is.na(fen_mt)) > 15,]

fen_chem <- read_csv('input/fen_feature_TWCD_mol_8-2-22.csv')
```

### Creating dissimilarity matrix

```{r}
fen_dist <- daisy(fen_mt_filt, metric = 'manhattan')
```

### Finding best k value

```{r cache=TRUE}
test_k(dist_matrix = fen_dist, k_init = 3, k_final = 10) 
```

Best value for k is **3**

### Getting representative features

```{r}
k <- 3

fen_clusters <- cluster_rep_features(dist_matrix = fen_dist, k = k)

rep_features_fen <- fen_clusters$representatives %>% 
  left_join(fen_chem, by = c('featureid' = 'Member')) %>% 
  ungroup()

write_csv(rep_features_fen, 'output/rep_features_fen.csv')

rep_features_fen_bnti <- fen_mt_filt %>% 
  rownames_to_column(var = 'featureid') %>% 
  mutate(featureid = as.double(featureid)) %>% 
  right_join(rep_features_fen, by = 'featureid') %>% 
  select(featureid, cluster, all_of(colnames(fen_mt_filt)))

write_csv(rep_features_fen_bnti, 'output/rep_features_fen_bnti.csv')
```







