---
title: "Metabolome paper - BNTI clusters correlations"
author: "Viviana Freire"
date: "`r Sys.Date()`"
output: html_document
---

# Loading libraries

```{r message=FALSE, warning=FALSE}
library(Hmisc)
library(tidyverse)
```

# Functions

```{r}
cluster_medians <- function(df, cluster, env_df){
  # Function to get the median value of cluster
  rownames(df) <- NULL
  df_tem <- df %>% 
    select(-cluster) %>% 
    column_to_rownames(var = 'featureid') %>% 
    select(all_of(env_df$SampleID))
  
  medians <- robustbase::colMedians(as.matrix(df_tem), na.rm = TRUE)
  
  medians <- as.data.frame(t(medians)) %>% 
    mutate(cluster = paste0('Cluster', cluster)) %>% 
    select(cluster, everything())
  
  return(medians)
}

env_cor <- function(env_var, cluster_df, env_df){
  # Function to correlate consensus cluster values with env data
  rownames(cluster_df) <- NULL
  cluster_df <- cluster_df[,c('cluster', env_df$SampleID)] %>% 
    column_to_rownames(var = 'cluster')
  
  env_data <- env_df[[env_var]]
  
  nclust <- nrow(cluster_df)
  
  corr <- Hmisc::rcorr(t(cluster_df), env_data, type = 'spearman')
  
  corr_r <- corr$r[1:nclust,'y']
  corr_p <- corr$P[1:nclust,'y']
  
  res <- tibble(cluster = rownames(cluster_df),
                env_variable = rep(env_var, nclust),
                rho = corr_r,
                pvalue = corr_p)
  
  res$pval_adj <- p.adjust(res$pvalue)
  
  return(res)
}

matrix_cor <- function(cluster_df, other_df, type){
  # Function to correlate with matrix data
  
  rownames(cluster_df) <- NULL
  cluster_df <- cluster_df %>% 
    column_to_rownames(var = 'cluster') %>% 
    select(all_of(colnames(other_df)))
  
  nclust <- nrow(cluster_df)
  
  # Keeping only features with more than 50% values
  
  if(type == 'bnti'){
    sel_rows <- rowSums(!is.na(other_df)) > ceil(ncol(other_df)/2)
  } else if(type == 'abundance'){
    sel_rows <- rowSums(other_df > 0) > ceil(ncol(other_df)/2)
  }
  
  other_df_filt <- other_df[sel_rows,]
  
  corr <- Hmisc::rcorr(t(cluster_df), t(other_df_filt), type = 'spearman')
  
  corr_r <- as.data.frame(corr$r) %>% 
    select(all_of(rownames(cluster_df))) %>% 
    rownames_to_column(var = 'Microbe') %>% 
    filter(!(Microbe %in% rownames(cluster_df))) %>% 
    pivot_longer(!Microbe, names_to = 'cluster', values_to = 'rho') %>% 
    drop_na()
  
  corr_p <- as.data.frame(corr$P) %>% 
    select(all_of(rownames(cluster_df))) %>% 
    rownames_to_column(var = 'Microbe') %>% 
    filter(!(Microbe %in% rownames(cluster_df))) %>% 
    pivot_longer(!Microbe, names_to = 'cluster', values_to = 'pvalue') %>% 
    drop_na()
  
  corr_p_adj <- as.data.frame(matrix(p.adjust(corr$P, method = 'fdr'),
                                     nrow = nrow(corr$P),
                                     dimnames = list(rownames(corr$P),
                                                     colnames(corr$P)))) %>% 
    select(all_of(rownames(cluster_df))) %>% 
    rownames_to_column(var = 'Microbe') %>% 
    filter(!(Microbe %in% rownames(cluster_df))) %>% 
    pivot_longer(!Microbe, names_to = 'cluster', values_to = 'p_adj') %>% 
    drop_na()
  
  res <- inner_join(corr_r, corr_p, by = c('Microbe', 'cluster')) %>% 
    inner_join(corr_p_adj, by = c('Microbe', 'cluster'))
  
  return(res)
}
```

# Data loading and preprocessing

## Microbial BNTI_feature data

```{r}
palsa_micro <- read_csv("input/bNTI_Feat_microbe_bulk_palsa_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1') 

bog_micro <- read_csv("input/bNTI_Feat_microbe_bulk_bog_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1')

fen_micro <- read_csv("input/bNTI_Feat_microbe_bulk_fen_bNTI_feature_by_samp_999rep.csv") %>% 
  column_to_rownames(var = '...1')
```

## Environmental data

```{r}
envs <- read_csv("input/peat_geodata_5-20-23.csv")

palsa_envs <- envs %>% 
  filter(Habitat == "Palsa") %>% 
  select(!c(Notes__.Biogeochemistry, Notes__.DepthInfo, 
            d13C_peat__, DepthMin__, Year, Samplenumber,
            Depth, Sampletype, DepthMax__, d15N_peat__, `...19`)) %>% 
  filter(SampleID %in% colnames(palsa_micro))

bog_envs <- envs %>% 
  filter(Habitat == "Bog") %>% 
  select(!c(Notes__.Biogeochemistry, Notes__.DepthInfo, 
            d13C_peat__, DepthMin__, Year, Samplenumber,
            Depth, Sampletype, DepthMax__, d15N_peat__, `...19`)) %>% 
  filter(SampleID %in% colnames(bog_micro))

fen_envs <- envs %>% 
  filter(Habitat == "Fen") %>% 
  select(!c(Notes__.Biogeochemistry, Notes__.DepthInfo, 
            d13C_peat__, DepthMin__, Year, Samplenumber,
            Depth, Sampletype, DepthMax__, d15N_peat__, `...19`)) %>% 
  filter(SampleID %in% colnames(fen_micro))

env_variables <- colnames(palsa_envs)[4:10]
```

## Metabolite BNTI_feature cluster data

```{r}
palsa_rep_values <- read_csv('output/rep_features_palsa_bnti.csv') %>% 
  select(featureid, cluster, colnames(palsa_micro)) %>% 
  split(.$cluster)

bog_rep_values <- read_csv('output/rep_features_bog_bnti.csv') %>% 
  select(featureid, cluster, colnames(bog_micro)) %>% 
  split(.$cluster)

fen_rep_values <- read_csv('output/rep_features_fen_bnti.csv') %>% 
  select(featureid, cluster, colnames(fen_micro)) %>% 
  split(.$cluster)
```

## CO2 data

```{r}
palsa_gas <- read_csv("input/palsa_matrix_3-1-22.csv") %>%
  select(!c(EMERGE_sampleID, SampleID_old__)) %>%
  select(SampleID, Month, Habitat, CH4.percent, CO2.percent) %>%
  #delete samples with NA
  na.omit()

bog_gas <- read_csv("input/bog_matrix_3-1-22.csv") %>%
  select(!c(EMERGE_sampleID, SampleID_old__)) %>%
  select(SampleID, Month, Habitat, pH_porewater, CH4.mM__, CO2.mM__) %>%
  #delete samples with NA
  na.omit()

fen_gas <- read_csv("input/fen_matrix_3-1-22.csv") %>%
  select(!c(EMERGE_sampleID, SampleID_old__)) %>%
  select(SampleID, Month, Habitat, pH_porewater, CH4.mM__, CO2.mM__) %>%
  #delete samples with NA
  na.omit()
```

## Microbial abundance

```{r}
mags_metadata <- read_csv('input/metadata_matched_MAGs_data_11-29-21.csv')

matched_names = setNames(mags_metadata$SampleID,
                         nm = mags_metadata$names_bnti)

otu_data <- read_csv("input/otu_normalized_df_3-4-22.csv") 

otu_table <- otu_data %>% 
  column_to_rownames(var = 'genome') %>% 
  select(all_of(mags_metadata$names_bnti)) %>% 
  rename_with(.fn = function(x){
    matched_names[x]
  })

otu_tax <- otu_data %>% 
  select(c(genome, Domain, Phylum, Class, Order, Family, Genus, Species)) %>% 
  rename(Microbe = genome)

otu_palsa <- otu_table %>% 
  select(all_of(colnames(palsa_micro)))

otu_bog <- otu_table %>% 
  select(all_of(colnames(bog_micro)))

otu_fen <- otu_table %>% 
  select(all_of(colnames(fen_micro)))
```

# Palsa correlations

## Getting consensus BNTI values per each cluster

```{r}
consensus_palsa <- imap(palsa_rep_values, ~cluster_medians(df = .x, 
                                                           cluster = .y,
                                                           env_df = palsa_envs))  
consensus_palsa_df <- do.call(rbind, consensus_palsa)
```

## Correlation with environmental variables

```{r}
corr_palsa_env <- map(env_variables, ~env_cor(env_var = .x,
                                              cluster_df = consensus_palsa_df,
                                              env_df = palsa_envs))

corr_palsa_env_df <- do.call(rbind, corr_palsa_env)
```

## Correlation with gas data

```{r}
corr_palsa_gas_df <- env_cor(env_var = 'CO2.percent',
                             cluster_df = consensus_palsa_df,
                             env_df = palsa_gas)
```

## Correlation with microbial BNTI

```{r}
corr_palsa_bnti_df <- matrix_cor(cluster_df = consensus_palsa_df,
                                 other_df = palsa_micro,
                                 type = 'bnti')
```

## Correlation with microbial abundances

```{r}
corr_palsa_abundance_df <- matrix_cor(cluster_df = consensus_palsa_df,
                                      other_df = otu_palsa,
                                      type = 'abundance')

corr_palsa_abundance_df_sig_taxa <- corr_palsa_abundance_df %>% 
  filter(p_adj < 0.05) %>% 
  left_join(otu_tax, by = 'Microbe')
```

## Saving correlation tables

```{r}
write_csv(corr_palsa_env_df, 'output/palsa_env_correlations.csv')
write_csv(corr_palsa_gas_df, 'output/palsa_gas_correlations.csv')
write_csv(corr_palsa_bnti_df, 'output/palsa_bnti-micro_correlations.csv')
write_csv(corr_palsa_abundance_df, 'output/palsa_abundance-micro_correlations.csv')
write_csv(corr_palsa_abundance_df_sig_taxa, 'output/palsa_sig_corr_taxa.csv')
```

# Bog correlations

## Getting consensus BNTI values per each cluster

```{r}
consensus_bog <- imap(bog_rep_values, ~cluster_medians(df = .x, 
                                                       cluster = .y,
                                                       env_df = bog_envs))  
consensus_bog_df <- do.call(rbind, consensus_bog)
```

## Correlation with environmental variables

```{r}
corr_bog_env <- map(env_variables, ~env_cor(env_var = .x,
                                            cluster_df = consensus_bog_df,
                                            env_df = bog_envs))

corr_bog_env_df <- do.call(rbind, corr_bog_env)
```

## Correlation with gas data

```{r}
corr_bog_gas <- map(c('CO2.mM__', 'CH4.mM__'), 
                    ~env_cor(env_var = .x,
                             cluster_df = consensus_bog_df,
                             env_df = bog_gas))

corr_bog_gas_df <- do.call(rbind, corr_bog_gas)
```

## Correlation with microbial BNTI

```{r}
corr_bog_bnti_df <- matrix_cor(cluster_df = consensus_bog_df,
                               other_df = bog_micro,
                               type = 'bnti')
```

## Correlation with microbial abundances

```{r}
corr_bog_abundance_df <- matrix_cor(cluster_df = consensus_bog_df,
                                    other_df = otu_bog,
                                    type = 'abundance')

corr_bog_abundance_df_sig_taxa <- corr_bog_abundance_df %>% 
  filter(p_adj < 0.05) %>% 
  left_join(otu_tax, by = 'Microbe')
```

## Saving correlation tables

```{r}
write_csv(corr_bog_env_df, 'output/bog_env_correlations.csv')
write_csv(corr_bog_gas_df, 'output/bog_gas_correlations.csv')
write_csv(corr_bog_bnti_df, 'output/bog_bnti-micro_correlations.csv')
write_csv(corr_bog_abundance_df, 'output/bog_abundance-micro_correlations.csv')
write_csv(corr_bog_abundance_df_sig_taxa, 'output/bog_sig_corr_taxa.csv')
```

# Fen correlations

## Getting consensus BNTI values per each cluster

```{r}
consensus_fen <- imap(fen_rep_values, ~cluster_medians(df = .x, 
                                                       cluster = .y,
                                                       env_df = fen_envs))  
consensus_fen_df <- do.call(rbind, consensus_fen)
```

## Correlation with environmental variables

```{r}
corr_fen_env <- map(env_variables, ~env_cor(env_var = .x,
                                            cluster_df = consensus_fen_df,
                                            env_df = fen_envs))

corr_fen_env_df <- do.call(rbind, corr_fen_env)
```

## Correlation with gas data

```{r}
corr_fen_gas <- map(c('CO2.mM__', 'CH4.mM__'), 
                    ~env_cor(env_var = .x,
                             cluster_df = consensus_fen_df,
                             env_df = fen_gas))

corr_fen_gas_df <- do.call(rbind, corr_fen_gas)
```

## Correlation with microbial BNTI

```{r}
corr_fen_bnti_df <- matrix_cor(cluster_df = consensus_fen_df,
                               other_df = fen_micro,
                               type = 'bnti')
```

## Correlation with microbial abundances

```{r}
corr_fen_abundance_df <- matrix_cor(cluster_df = consensus_fen_df,
                                    other_df = otu_fen,
                                    type = 'abundance')

corr_fen_abundance_df_sig_taxa <- corr_fen_abundance_df %>% 
  filter(p_adj < 0.05) %>% 
  left_join(otu_tax, by = 'Microbe')
```

## Saving correlation tables

```{r}
write_csv(corr_fen_env_df, 'output/fen_env_correlations.csv')
write_csv(corr_fen_gas_df, 'output/fen_gas_correlations.csv')
write_csv(corr_fen_bnti_df, 'output/fen_bnti-micro_correlations.csv')
write_csv(corr_fen_abundance_df, 'output/fen_abundance-micro_correlations.csv')
write_csv(corr_fen_abundance_df_sig_taxa, 'output/fen_sig_corr_taxa.csv')
```


